Lattice Boltzmann simulations of magnetically responsive\\ anisotropic particle stabilized emulsions

### 2025-02-19 chapters/introduction.tex
Changes 1. Added more analysis to application of magnetic fields during formation chapter 2. Added references to fix some (not all) literature reference issues 3. Added some detail in analysis techniques section
Changes 1. Adjusted introduction for flow and fixed references

\begin{itemize}
    \item What is a porous material? 
    \begin{itemize}
        \item What makes them different to non-porous materials?
        \item Why do their differences make them industrially relevant?
        \item What are some synthesis techniques that exist?
        \item What are weaknesses with those synthesis techniques?
    \end{itemize}
    \item What is a particle stabilized emulsion
    \begin{itemize}
        \item Why are they interesting for porous material applications?
        \item What is a bijel?
        \item How are bijels relevant to porous materials?
        \item Fabrication techniques that are proposed which allow large scale bijel fabrication
        \item microstructure control in those applications
    \end{itemize}
    \item Purpose of study
    \begin{itemize}
        \item Three aims of the study 
        \item significance of study
        \item assumptions and limitations of the study
    \end{itemize}
    \item Summary of motivation, purpose and research aims
\end{itemize}

@@

\section{Introduction}

Porous materials are defined as materials that contain pores or voids within their structure where no material is present. 
Homogeneous materials in contrast have a uniform structure with no pores. These pores vary in shape and distribution, with materials 
being categorized by pore size per IUPAC as $(L)$ as microporous $(L < 2nm)$, mesoporous $(2nm < L < 50nm)$ and macroporous materials 
$(50nm < L)$. \textcolor{blue}{https://doi.org/10.1351/pac198557040603} The presence of these pores increases the surface area to volume 
ratio of porous materials compared to their homogeneous counterparts. A common experimental technique used to estimate the surface area 
of a porous material derived from Brunauer-Emmett-Teller(BET) theory shows that porous materials can have surface areas more than a 
hundred times larger than homogenous materials, dependent upon the pore size and pore size distribution. \cite{shimizu_surface_2022}

In addition to the pore size and its distribution within the material, the arrangement of pores within the material contribute to the 
performance of the synthesized material, commonly measured using the tortuosity. Tortuosity here is defined as the path length between 
two points divided by the euclidean length between those points. A tortuosity larger than one implies that the path length is longer 
than the euclidean length, and vice versa if the tortuosity if less than one. Highlighting the confluence of these factors are the 
large variety of tailored microstructures for various applications. Each of the specified pore length scales have their unique uses, 
such as microporous materials being used as zeolites and photonic bandgap materials while macroporous materials facilitate greater 
transport properties, suitable in applications such as battery electrodes and drug delivery systems. \cite{chen_tortuosity_2020, 
ebner_tortuosity_2014}

These range from Metal Organic Framework(MOF) enhanced gas adsorption, \textcolor{blue}{https://doi.org/10.1002/admi.201400040} 
This increased surface area to volume ratio enhances the properties of the base material in applications such as catalysis, 
battery electrodes and drug delivery systems among many more making them valuable microstructures. \
cite{cha_bicontinuous_2019, samdani_bicontinuous_2017, thorson_bijel-templated_2019}

@@

Stimuli response based on pH, external fields and temperature have been used in past works when looking at modifying the microstructure of 
emulsion droplets for enhanced oil recovery, pharmaceutical and cell adhesion. \cite{haase_nanoparticle_2011, tham_magnetophoresis_2021, 
cui_stabilizing_2013, manfredini_limonene--water_2021}  It was shown that external fields can be used to move, control emulsion stability 
and elongate emulsion droplets. \cite{tham_magnetophoresis_2021, cui_stabilizing_2013} 
\textcolor{blue}{https://pubs.acs.org/doi/10.1021/la047691n} Bijels stabilized spherical particles under a magnetic field showed no 
meaningful microstructure changes. \cite{kim_bijels_2010} Bijels stabilized with spherical particles under an electric field showed 
more promise. \cite{carmack_tuning_2018} However, magnetic fields offer more targeted stimuli response, allowing even weak fields to 
incite a response. Magnetic fields non-interaction in many situations offer advantages to many of the desired uses of bijels within 
pharmaceutical or bioengineering applications. \cite{vanoli_bijels_2022, thorson_bijel-templated_2019, thorson_composite_2018} These 
same applications also have a need for in-situ microstructure modification that would enable tunable drug delivery rates, separations or
 catalyst efficiency and permeability improvements in the system. Therefore, identifying schemes that allow modification of the pore size 
 and tortuosity  without synthesis of a new material would enhance the use case of bijels.

Anisotropic particles at interfaces have shown self assembly due to the presence of multi-polar pressure fields caused by interface deformation. 
Self assembly has also been shown for ellipsoidal particles at interfaces under magnetic fields. Ellipsoidal particles have also been shown to 
exhibit magnetic field controlled orientations at interfaces, as Bresme and Faraudo showed. \cite{bresme_orientational_2007, davies_interface_2014}
 Under magnetic fields, assemblies of multiple ellipsoidal particles have been shown to have energy minima at unexpected locations. 
 \cite{newton_influence_2014, newton_capillary_2018} Thus ellipsoidal particles at interfaces, whose orientations and packing can be 
 controlled through external fields, offer a potential means to control the microstructure of bijels.

Bijels present a bottom up synthesis strategy for porous material templates for hard and soft material applications in a scalable and 
continuous method using STrIPS and other techniques. However, the microstructure control these techniques offer are intrinsically linked 
to the casting mixture composition. Stimuli response offers a way to modify the microstructure of bijels while still maintaining the 
various form factors these continuous synthesis techniques can output. While stimuli response has been explored in the past using bijels 
stabilized with spherical particles with little success, anisotropic particles are poised to allow microstructure modification of bijels 
upon application of an external field. This technique also has the potential to allow for post synthesis microstructure modification. 
Therefore, this work seeks to identify a means to control the microstructure of a bijel stabilized with magnetically responsive 
ellipsoidal particles using magnetic fields.

% Bijels have proposed uses in many of these with the microstructure of the material being shown to affect material properties. \cite{cha_bicontinuous_2019, khan_nanostructured_2022, vanoli_bijels_2022}

% Magnetic fields offer several advantages over currently utilized techniques such as electric fields and shear. \cite{cui_stabilizing_2013, mulligan_deformation_2011} These include greater specificity of response as most materials are unresponsive to magnetic stimuli as well as even weak fields being sufficient to incite responses in many cases.

% this adds value to many of the desired uses of bijels within pharmaceutical or bioengineering applications. \cite{vanoli_bijels_2022, thorson_bijel-templated_2019, thorson_composite_2018}

\textcolor{blue}{Add commentary about past results with bijels and why we use anisotropic particles in bijels. Changes this makes} Therefore, 
identification of a scheme which allows for tunability of the microstructure, independent of the composition of the casting mixture and post 
synthesis is the goal of this work.

@@

This work seeks to identify means to control the bijel microstructure independent of the casting mixture composition by utilizing magnetic 
stimuli on bijels stabilized with magnetically responsive ellipsoidal particles. The proposed mechanism of microstructure modification is 
driven through reorientation of the particles to the applied magnetic field altering how the particles pack on the interface. By changing the 
interfacial packing of the particles, we change when the particles jam, which in turn changes the jamming point of the bijel. This mechanism is 
predicted to be effective during and after synthesis. The processability of bijels fabricated using this technique will also be investigated to 
ascertain how changes in particle packing affect the rheology of the material. An overview of three major research aims will be presented below.

@@

The jamming point of a bijel controls the obtained microstructure, as shown by previous work comparing the length scale of bijels synthesized 
with different particle volume fractions. It has also been shown how magnetic fields can be used to self assemble ansisotropic particles at 
interfaces and how they can tilt out of the interface. \cite{davies_interface_2014, davies_assembling_2014} On a curved interface with many 
other particles, it is unknown how multiple particles orienting to a field and having regular structure will affect when and how the bijel 
will jam. \cite{bresme_orientational_2007, davies_interface_2014}

To assess the obtained microstructure upon application of a field during formation of the bijel, homogenous mixtures containing spherical and 
ellipsoidal particles will be subjected to field strengths above and below the critical field strength when particle orientations to the 
interface are controlled by the applied field strength, as calculated by Bresme and Faraudo for the particle used. This selection of fields 
will allow insight into the controlling mechanisms behind microstructural changes observed, allowing verification of the hypothesis and 
ascertainment where the most control over the microstructure can be exercised.

The average and directional domain sizes in addition to their relation with tortuosity will provide windows into the microstructural properties 
of the resulting bijel. The nematic order parameter and average interfacial angle will be used to investigate how the application of the 
field affects the particle ordering to each other, and to the interface. Finally, the radial distribution function will be used to assess 
if the orientational changes expected upon field application, affects the packing of the particles on the interface.

@@

Particles on the surface of emulsion droplets have been shown to unjam and rejam into new, stable microstructures upon application of 
stimuli. \cite{cui_stabilizing_2013} Two techniques of microstructure change are proposed. The first technique is that the adsorption 
energy is so strong that the interface is dragged along with the particles until they rejam in their new location and orientation to the 
field. The second technique is that the particles tilt in place out of the interface, causing domain coarsening before the particles jam 
in their new positions and orientations. The mechanism, derived microstructure change and timescales will be assessed in this aim.

@@

In many of the fabrication processes described above, the rheology of the bijel casting mixture is essential. 
\cite{haase_continuous_2015, cai_bijels_2017, amirfattahi_fabrication_2024} Ching showed that rheologically, bijels are 
colloidal glasses percolating in 3D space. \cite{ching_bijel_2022} First, a shear capillary number is defined as 
$Ca_s = \frac{\eta_{f} \dot{\gamma} L_{1}}{\sigma}$ where $\dot{\gamma} = \frac{u_{LE}}{L_x}$ is the strain rate and 
$L_1$ is the average domain size. \cite{frijters_effects_2012, yang_capillary_2022} Lower capillary numbers in the range of
 $ 10^{-7} \geq Ca_s \leq 10^{-5}$ will be utilized. 
% which corresponds to a $Ma << 0.01$ to accommodate the hydrodynamic model utilized.

For this study, the yield stress and viscosity of the bijels will be explored as a function of the initial microstructure under 
various $Ca_s$. Four templates with different processing history will be utilized to assess the impact of initial microstructure on 
the obtained rheology. These are, $\bar{B} = 0$, $\bar{B} = 1$, $\bar{B} = 0 \rightarrow \bar{B} = 1$ and $\bar{B} = 1 \rightarrow \bar{B} = 0$. 
To explain any observed differences in the the yield stress and viscosity, the time evolution of the domain size, average interface angle, and 
nematic order parameter will be characterized and compared to the observed yield stress and viscosity. Additionally, the presence of shear 
banding will be visually confirmed.

>>>>

% \begin{itemize}
%     \item What is a porous material? 
%     \begin{itemize}
%         \item What makes them different to non-porous materials?
%         \item Why do their differences make them industrially relevant?
%         \item What are some synthesis techniques that exist?
%         \item What are weaknesses with those synthesis techniques?
%     \end{itemize}
%     \item What is a particle stabilized emulsion
%     \begin{itemize}
%         \item Why are they interesting for porous material applications?
%         \item What is a bijel?
%         \item How are bijels relevant to porous materials?
%         \item Fabrication techniques that are proposed which allow large scale bijel fabrication
%         \item microstructure control in those applications
%     \end{itemize}
%     \item Purpose of study
%     \begin{itemize}
%         \item Three aims of the study 
%         \item significance of study
%         \item assumptions and limitations of the study
%     \end{itemize}
%     \item Summary of motivation, purpose and research aims
% \end{itemize}

@@

\section{Introduction}
\label{section:introduction}

Porous materials are defined as materials that contain pores or voids within their structure where no material is present. 
Homogeneous materials in contrast have a uniform structure with no pores. These pores vary in shape and distribution, with materials 
being categorized by pore size per IUPAC as $(L)$ as microporous $(L < 2nm)$, mesoporous $(2nm < L < 50nm)$ and macroporous materials 
$(50nm < L)$. \cite{sing_reporting_1985} The presence of these pores increases the surface area to volume 
ratio of porous materials compared to their homogeneous counterparts. A common experimental technique used to estimate the surface area 
of a porous material derived from Brunauer-Emmett-Teller(BET) theory shows that porous materials can have surface areas more than a 
hundred times larger than homogenous materials, dependent upon the pore size and pore size distribution. \cite{shimizu_surface_2022}

In addition to the pore size and its distribution within the material, the arrangement of pores within the material contribute to the 
performance of the synthesized material, commonly measured using the tortuosity. Tortuosity here is defined as the path length between 
two points divided by the euclidean length between those points. A tortuosity larger than one implies that the path length is longer 
than the euclidean length, and vice versa if the tortuosity if less than one. Highlighting the confluence of these factors are the 
large variety of tailored microstructures for various applications. Each of the specified pore length scales have their unique uses, 
such as microporous materials being used as zeolites and photonic bandgap materials while macroporous materials facilitate greater 
transport properties, suitable in applications such as battery electrodes and drug delivery systems. \cite{chen_tortuosity_2020, 
ebner_tortuosity_2014} This increased surface area to volume ratio enhances the properties of the base material in applications such 
as catalysis, battery electrodes and drug delivery systems among many more making them valuable microstructures. 
\cite{cha_bicontinuous_2019, samdani_bicontinuous_2017, thorson_bijel-templated_2019, zhao_highly_2014}

@@

Stimuli response based on pH, external fields and temperature have been used in past works when looking at modifying the microstructure of 
emulsion droplets for enhanced oil recovery, pharmaceutical and cell adhesion. \cite{haase_nanoparticle_2011, tham_magnetophoresis_2021, 
cui_stabilizing_2013, manfredini_limonene--water_2021}  It was shown that external fields can be used to move, control emulsion stability 
and elongate emulsion droplets. \cite{tham_magnetophoresis_2021, cui_stabilizing_2013, melle_pickering_2005} Bijels stabilized spherical 
particles under a magnetic field showed no meaningful microstructure changes. \cite{kim_bijels_2010} Bijels stabilized with spherical 
particles under an electric field showed more promise. \cite{carmack_tuning_2018} However, magnetic fields offer targeted stimuli 
response, allowing even weak fields to incite a response. Magnetic fields non-interaction in many situations offer advantages to many of 
the desired uses of bijels within pharmaceutical or bioengineering applications. \cite{vanoli_bijels_2022, thorson_bijel-templated_2019, thorson_composite_2018} 
These applications can also take advantage of in-situ microstructure modification that would enable tunable drug delivery rates, separations or
catalyst efficiency and permeability improvements in the system. Therefore, identifying schemes that allow modification of the pore size 
and tortuosity without synthesis of a new material would enhance the use case of bijels.

Anisotropic particles at interfaces have shown self assembly due to the presence of multi-polar pressure fields caused by interface deformation. 
Self assembly has also been shown for ellipsoidal particles at interfaces under magnetic fields. Ellipsoidal particles have also been shown to 
exhibit magnetic field controlled orientations at interfaces, as Bresme and Faraudo showed. \cite{bresme_orientational_2007, davies_interface_2014}
Under magnetic fields, assemblies of multiple ellipsoidal particles have been shown to have energy minima at orientations different from isolated particles,
meaning that the interface coverage of interfaces decorated with ellipsoidal particles can be varied using stimuli response. 
\cite{newton_influence_2014, newton_capillary_2018} Thus ellipsoidal particles at interfaces, whose orientations and packing can be 
controlled through external fields, offer a potential means to control the microstructure of bijels.

Bijels present a bottom up synthesis strategy for porous material templates for hard and soft material applications in a scalable and 
continuous method using STrIPS and other techniques. However, the microstructure control these techniques offer are intrinsically linked 
to the casting mixture composition. Stimuli response offers a way to modify the microstructure of bijels while still maintaining the 
various form factors these continuous synthesis techniques can output. Magnetic stimuli response has been explored in the past using bijels 
stabilized with spherical particles with little success. \cite{kim_bijels_2010}

With this background, we propose anisotropic particles as a technique to enable microstructure modification of bijels 
upon application of an external field during and after fabrication. This work seeks to identify means to control the bijel microstructure independent 
of the casting mixture composition by utilizing magnetic stimuli on bijels stabilized with magnetically responsive ellipsoidal particles. 
The proposed mechanism of microstructure modification is driven through reorientation of the particles to the applied magnetic field altering 
how the particles pack on the interface. By changing the interfacial packing of the particles, we change when the particles jam, which in turn changes 
the jamming point of the bijel. This mechanism is predicted to be effective during and after synthesis. The processability of bijels fabricated using 
this technique will also be investigated to ascertain how changes in particle packing affect the rheology of the material. An overview of three major 
research aims will be presented below.

@@

@@

The jamming point of a bijel controls the obtained microstructure, as shown by previous work comparing the length scale of bijels synthesized 
with different particle volume fractions and particle sizes. \cite{jansen_bijels_2011, reeves_particle-size_2015} It has also been shown how magnetic fields 
can be used to enable self assembly of ansisotropic particles at interfaces and control the interfacial angle between the particle axis of symmetry and the interface. 
\cite{davies_interface_2014, davies_assembling_2014}
On a curved interface with many other particles, it is unknown how multiple particles orienting to a field and having regular structure will affect when and how the bijel 
will jam. \cite{bresme_orientational_2007, davies_interface_2014}

To assess the obtained microstructure upon application of a field during formation of the bijel, homogenous mixtures containing spherical and 
ellipsoidal particles will be subjected to field strengths above and below the critical field strength when particle orientations to the 
interface are controlled by the applied field strength, as calculated by Bresme and Faraudo for the particle aspect ratio. \cite{bresme_orientational_2007, davies_interface_201} This selection of fields 
will allow insight into the controlling mechanisms behind microstructural changes observed, allowing verification of the hypothesis and 
ascertainment where the most control over the microstructure can be exercised.

The average and directional domain sizes in addition to their relation with tortuosity will allow characterization of the microstructural properties 
of the resulting bijel. The nematic order parameter and average interfacial angle will be used to investigate how the application of the 
field affects the particle ordering to each other, and to the interface. Finally, the radial distribution function will be used to assess 
if the orientational changes expected upon field application, affects the packing of the particles on the interface.

@@

Particles on the surface of emulsion droplets have been shown to unjam and rejam into new, stable microstructures upon application of 
stimuli. \cite{cui_stabilizing_2013} Magnetic fields applied to bijels stabilized with ellipsoidal particles will facilitate unjamming and rejamming of
the particle monolayer, resulting in microstructure modification of the bijel. Due to local changes in the state of the particles, specifically the 
interfacial ordering in a specific direction and arrangement of particles, we expect that there will be microstructure modification even when the 
magnetic field is switched off. Two techniques of microstructure change are proposed. The first technique is that the adsorption 
energy is so strong that the interface is dragged along with the particles until they rejam in their new location and orientation to the 
field. The second technique is that the particles tilt in place out of the interface, causing domain coarsening before the particles jam 
in their new positions and orientations. The mechanism, derived microstructure change and timescales will be assessed in this aim.

@@

In many of the fabrication processes described above, the rheology of the bijel casting mixture is essential. 
\cite{haase_continuous_2015, cai_bijels_2017, amirfattahi_fabrication_2024} Ching showed that rheologically, bijels are 
colloidal glasses percolating in 3D space. \cite{ching_bijel_2022} A shear capillary number, $Ca_s$, is defined that correlates the 
applied strain rate to the capillary forces derived from surface tension. \cite{frijters_effects_2012, yang_capillary_2022} We are interested 
in identifying the effect of the initial microstructure and an applied field on the rheological properties of the bijel.

% $Ca_s = \frac{\eta_{f} \dot{\gamma} L_{1}}{\sigma}$ where $\dot{\gamma} = \frac{u_{LE}}{L_x}$ is the strain rate and 
% $L_1$ is the average domain size. \cite{frijters_effects_2012, yang_capillary_2022} Lower capillary numbers in the range of
%  $ 10^{-7} \geq Ca_s \leq 10^{-5}$ will be utilized. 
% which corresponds to a $Ma << 0.01$ to accommodate the hydrodynamic model utilized.

For this study, the yield stress and viscosity of the bijels will be explored as a function of the initial microstructure under 
various $Ca_s$. The initial microstructure will be derived from bijels stabilized with ellipsoidal particles made with no magnetic field 
and with a field strength equal to the surface tension energy. We then apply constant shear defined with $Ca_s$ to determine the yield stress and non-newtonian
characteristics of the bijel. Previous investigations have revealed that the ordering of particles in the direction of shear change
the rheological behavior significantly compared to unordered systems. Additionally, the domains of the bijel may create additional mechanisms for 
rheological behavior modification which have not been probed in detail yet.

We will then identify the complex rheology of the bijels, calculating the loss and storage modulus of the bijel microstructures detailed earlier
as a function of the oscillation of shear. The loss and storage modulus detail the viscous and elastic behavior of the bijel respectively, meaning 
that the gel point of the bijel can be established as the point when the storage modulus exceeds the loss modulus. We will pay special attention to the 
state of the particles in all simulations, correlating changes in the properties of the particle to observed rheological events. We will also identify if shear banding
is observed.
